# Data Cleaning

In this section, we clean and prepare the dataset for analysis. Data cleaning is a critical step in any workflow, as real-world data is often incomplete or inconsistent.

## Clean and Transform the Data

We will:

- select relevant variables  
- convert categorical variables to factors  
- create a new feature  
- remove missing values

```{r}
#| output: false
#| echo: false

library(tidyverse)
library(palmerpenguins)
df_raw <- penguins
str(df_raw)
head(df_raw)
summary(df_raw)
```

```{r}
df <- df_raw %>%
  select(species, island, bill_length_mm, bill_depth_mm, flipper_length_mm,
         body_mass_g, sex) %>%
  mutate(
    species = as.factor(species),
    island = as.factor(island),
    sex = as.factor(sex),
    large_body = ifelse(body_mass_g > median(body_mass_g, na.rm = TRUE), 1, 0)
  ) %>%
  filter(
    !is.na(bill_length_mm),
    !is.na(bill_depth_mm),
    !is.na(flipper_length_mm),
    !is.na(body_mass_g),
    !is.na(sex)
  )
```

## Inspect the Cleaned Data

```{r}
str(df)
summary(df)
```

## Compare Before and After

```{r}
nrow(df_raw)
nrow(df)
```

This comparison shows how many observations were removed during cleaning, reinforcing that data preparation can significantly impact your dataset.
