# Exploratory Data Analysis

```{r}
#| output: false
#| echo: false

library(tidyverse)
library(palmerpenguins)
df_raw <- penguins
str(df_raw)
head(df_raw)
summary(df_raw)
df <- df_raw %>%
  select(species, island, bill_length_mm, bill_depth_mm, flipper_length_mm,
         body_mass_g, sex) %>%
  mutate(
    species = as.factor(species),
    island = as.factor(island),
    sex = as.factor(sex),
    large_body = ifelse(body_mass_g > median(body_mass_g, na.rm = TRUE), 1, 0)
  ) %>%
  filter(
    !is.na(bill_length_mm),
    !is.na(bill_depth_mm),
    !is.na(flipper_length_mm),
    !is.na(body_mass_g),
    !is.na(sex)
  )
```

In this section, we begin exploring the cleaned dataset. Exploratory Data Analysis (EDA) helps us understand the structure of the data, identify patterns, and guide our modeling decisions.

EDA is an important step because it allows us to:
- understand distributions of variables  
- identify relationships between variables  
- detect potential outliers or unusual values  
- build intuition before fitting models  

## Distribution of Body Mass

We start by examining the distribution of body mass to understand its overall shape.

```{r}
ggplot(df, aes(x = body_mass_g)) +
  geom_histogram(bins = 12) +
  labs(title = "Distribution of Body Mass")
```

## Body Mass vs Flipper Length

Next, we explore the relationship between flipper length and body mass.

```{r}
ggplot(df, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  labs(title = "Body Mass vs Flipper Length")
```

## Body Mass vs Bill Length

We also examine how bill length relates to body mass.

```{r}
ggplot(df, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point() +
  labs(title = "Body Mass vs Bill Length")
```

## Body Mass by Species

Finally, we compare body mass across different penguin species.

```{r}
ggplot(df, aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  labs(title = "Body Mass by Species")
```

## Key Takeaways

From these plots, we can begin to see patterns in the data:

- body mass has a clear distribution  
- some variables appear to have relationships with body mass  
- differences between species may be important  

These insights will guide how we build and interpret our models in the next section.
