# Linear Regression

```{r}
#| output: false
#| echo: false

library(tidyverse)
library(palmerpenguins)
df_raw <- penguins
str(df_raw)
head(df_raw)
summary(df_raw)
df <- df_raw %>%
  select(species, island, bill_length_mm, bill_depth_mm, flipper_length_mm,
         body_mass_g, sex) %>%
  mutate(
    species = as.factor(species),
    island = as.factor(island),
    sex = as.factor(sex),
    large_body = ifelse(body_mass_g > median(body_mass_g, na.rm = TRUE), 1, 0)
  ) %>%
  filter(
    !is.na(bill_length_mm),
    !is.na(bill_depth_mm),
    !is.na(flipper_length_mm),
    !is.na(body_mass_g),
    !is.na(sex)
  )
```

Linear regression is a statistical method used to model the relationship between a continuous outcome and one or more predictor variables. It fits a line (or plane) that best describes how the outcome changes as the predictors change, allowing us to interpret relationships and make predictions.

## Fit the Model

We will predict body mass using several body measurements.

```{r}
lm_model <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = df)
```

## Inspect the Model

```{r}
summary(lm_model)
coef(lm_model)
```

- `summary()` provides overall model information and statistical significance.
- `coef()` shows the estimated effect of each predictor.

## Visualize the Relationship

To build intuition, we can visualize one predictor against the outcome and overlay the fitted regression line.

```{r}
ggplot(df, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Linear Regression: Body Mass vs Flipper Length")
```

## Key Takeaways

- Linear regression helps quantify relationships between variables.
- Coefficients indicate how the outcome changes with each predictor.
- Visualization is useful for understanding model fit and interpreting results.
