# Logistic Regression

```{r}
#| output: false
#| echo: false

library(tidyverse)
library(palmerpenguins)
df_raw <- penguins
str(df_raw)
head(df_raw)
summary(df_raw)
df <- df_raw %>%
  select(species, island, bill_length_mm, bill_depth_mm, flipper_length_mm,
         body_mass_g, sex) %>%
  mutate(
    species = as.factor(species),
    island = as.factor(island),
    sex = as.factor(sex),
    large_body = ifelse(body_mass_g > median(body_mass_g, na.rm = TRUE), 1, 0)
  ) %>%
  filter(
    !is.na(bill_length_mm),
    !is.na(bill_depth_mm),
    !is.na(flipper_length_mm),
    !is.na(body_mass_g),
    !is.na(sex)
  )
lm_model <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = df)
```

Logistic regression is a modeling technique used when the outcome variable is **binary**, meaning it has two possible values (such as 0/1, yes/no, or success/failure). Instead of predicting a continuous value, logistic regression estimates the **probability** that an observation belongs to a particular class.

## Create Train and Test Sets

As before, we split the data so that we can evaluate the model on unseen data.

```{r}
set.seed(123)

train_index_cls <- sample(seq_len(nrow(df)), size = floor(0.8 * nrow(df)))
train_cls <- df[train_index_cls, ]
test_cls  <- df[-train_index_cls, ]
```

In this example, the outcome variable is:

- `large_body = 1` → above-median body mass  
- `large_body = 0` → below-median body mass  

---

## Fit the Logistic Regression Model

```{r}
glm_model <- glm(
  large_body ~ bill_length_mm + bill_depth_mm + flipper_length_mm,
  data = train_cls,
  family = "binomial"
)
```

- `glm()` is used for generalized linear models  
- `family = "binomial"` specifies logistic regression  

---

## Inspect the Model

```{r}
summary(glm_model)
```

- Coefficients represent how each predictor affects the **log-odds** of the outcome  
- Positive values increase the probability of class 1  
- Negative values decrease the probability of class 1  

---

## Predict Probabilities

```{r}
prob_preds <- predict(glm_model, newdata = test_cls, type = "response")
```

- Predictions are probabilities between 0 and 1  
- These represent the likelihood of belonging to class 1  

---

## Convert Probabilities to Classes

```{r}
class_preds <- ifelse(prob_preds >= 0.5, 1, 0)
```

- A threshold (here, 0.5) is used to assign class labels  
- This converts probabilities into predictions  

---

## Evaluate the Model

```{r}
mean(class_preds == test_cls$large_body)
```

- This computes **accuracy**, the proportion of correct predictions  

```{r}
table(actual = test_cls$large_body, predicted = class_preds)
```

- This is a **confusion matrix**  
- It shows how many predictions were correct vs incorrect  

---

## Key Takeaways

- Logistic regression is used for **binary outcomes**  
- The model predicts probabilities, not just classes  
- A threshold is used to convert probabilities into predictions  
- Evaluation should be done on a **test dataset**  

> Like linear regression, the goal is not just to fit the data,  
> but to make reliable predictions on new observations.
