# Setup

In this section, we will set up our working environment and organize our project. A well-structured project is the foundation of a reproducible workflow.

## Create a Project

First, create an RStudio Project by selecting `File` > `New Project...`.

Then, donwload the code file and place in your project directory: [R Code](code/script.R)

## Load Required Packages

We will use a small number of packages throughout this workshop:

- **tidyverse** for data manipulation and visualization  
- **palmerpenguins** for our example dataset  

If you do not already have these installed, run:

```r
install.packages(c("tidyverse", "palmerpenguins"))
```

Now load the packages:

```{r}
#| output: false

library(tidyverse)
library(palmerpenguins)
```

## Set a Seed for Reproducibility

Setting a seed ensures that any random operations (such as splitting data into training and testing sets) produce the same results each time the code is run.

```{r}
set.seed(123)
```

## Project Organization

Before writing any code, it is important to organize your project.

In your RStudio Project, manually create the following folders:

- `data/` → for raw and cleaned datasets  
- `outputs/` → for results, plots, and exported files  

This structure helps keep your work organized and makes it easier to rerun or share your analysis.

## Create and Save the Dataset {.unnumbered}

We will use the **penguins** dataset for this workshop.

To simulate working with an external dataset, we will write it to a CSV file inside the `data/` folder. This allows us to practice reading data from a file, which is how most real-world workflows begin.

```r
example_path <- "data/penguins.csv"

if (!file.exists(example_path)) {
  write_csv(penguins, example_path)
}
```

## Key Takeaway

A reproducible workflow starts with:

- clearly organized folders  
- consistent file paths  
- explicit package usage  
- and code that can be rerun from start to finish  

These small habits make a big difference as your projects grow in complexity.
