# Training and Testing

```{r}
#| output: false
#| echo: false

library(tidyverse)
library(palmerpenguins)
df_raw <- penguins
str(df_raw)
head(df_raw)
summary(df_raw)
df <- df_raw %>%
  select(species, island, bill_length_mm, bill_depth_mm, flipper_length_mm,
         body_mass_g, sex) %>%
  mutate(
    species = as.factor(species),
    island = as.factor(island),
    sex = as.factor(sex),
    large_body = ifelse(body_mass_g > median(body_mass_g, na.rm = TRUE), 1, 0)
  ) %>%
  filter(
    !is.na(bill_length_mm),
    !is.na(bill_depth_mm),
    !is.na(flipper_length_mm),
    !is.na(body_mass_g),
    !is.na(sex)
  )
lm_model <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = df)
```

When building models, it is important to evaluate how well they perform on **new, unseen data**. If we only evaluate a model on the same data it was trained on, we can get overly optimistic results.

## Why Split the Data?

If we train and test on the same dataset:

- the model has already “seen” the data  
- it can appear more accurate than it really is  
- we cannot tell how well it will perform in practice  

To address this, we split the data into:

- **Training set** → used to fit the model  
- **Test set** → used to evaluate performance  

---

## Create Train and Test Sets

```{r}
set.seed(123)

train_index <- sample(seq_len(nrow(df)), size = floor(0.8 * nrow(df)))
train <- df[train_index, ]
test  <- df[-train_index, ]
```

- 80% of the data is used for training  
- 20% is held out for testing  

---

## Train the Model

```{r}
lm_train <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = train)
```

The model is fit **only on the training data**.

---

## The Wrong Way: Evaluate on Training Data

```{r}
train_preds_wrong <- predict(lm_train, newdata = train)
sqrt(mean((train$body_mass_g - train_preds_wrong)^2))
```

- This often produces a low error  
- But it is misleading because the model has already seen this data  

> This is like studying using the exact answers to the test.

---

## The Correct Way: Evaluate on Test Data

```{r}
test_preds <- predict(lm_train, newdata = test)
sqrt(mean((test$body_mass_g - test_preds)^2))
```

- This gives a more realistic estimate of performance  
- It reflects how the model behaves on new data  

---

## Visualizing Predictions

```{r}
eval_df <- test
eval_df$predicted_body_mass_g <- test_preds

ggplot(eval_df, aes(x = body_mass_g, y = predicted_body_mass_g)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Test Set Predictions")
```

- Points close to the dashed line indicate good predictions  
- Larger deviations indicate prediction errors  

---

## Key Takeaways

- Always evaluate models on **data they have not seen**  
- Training error is often misleading  
- Test error provides a better estimate of real-world performance  

> A good model is not one that fits the training data perfectly,  
> but one that generalizes well to new data.
